Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-08 paper/sec-analysis.tex
more work polishing the text
Changes

@@

In this analysis we adopt
a $b$-tagging strategy along the lines
of current ATLAS performance~\cite{Aad:2013gja,Aad:2015ydr},
though differences with respect to
the corresponding CMS case~\cite{Khachatryan:2011wq,Chatrchyan:2012jua}
do not modify qualitatively our results.
%
For each type of jet definition used, a different
$b$-tagging strategy is adopted:

@@

Once a Higgs boson candidate has been identified,
its invariant mass is required to lie within a fixed window
around the nominal Higgs boson mass $m_h\simeq 125$
GeV~\cite{Aad:2015zhl}.
%
In particular we require the condition
\be
\label{higgsmasswindow}
|m_{h,j} - 125| < 40~{\rm GeV} \, ,\, j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of each of the two reconstructed  Higgs candidates.
%
This cut is substantially looser as the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}: the motivation
for this choice is that the MVA will take care of the optimisation of this,
as well as of other cuts that are not explicitly performed.
%
Only events where the two Higgs candidates satisfy
Eq.~(\ref{higgsmasswindow}) are classified as signal events.
%

@@

 This has been achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with higher significance turns out to be the boosted,
followed by the intermediate and the resolved ones, the two latter
with similar significance.
%
Therefore, for the final analysis, we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is assigned to
this category, else we check if it suits the intermediate
requirements, and failing also this, the event is attempted to
be classified in the
resolved category.
%
The resulting exclusive event samples are then separately processed
through the MVA, allowing a consistent combination
of the significance of the three event categories.

\subsection{Motivation for basic kinematical cuts}

We now motivate the values of
kinematical cuts applied to the different categories, 
comparing representative kinematic distributions between
signal and background events.
%
First to all we present results without PU, and in the next
subsection we discuss what is the impact of PU
on the description of the various kinematical
distributions.
%
In the following, all
distributions will be normalized to their total integral.

%
To begin with,
in Fig.~\ref{fig:cutplots1} we show
the $p_T$ distribution of the 
of the
  leading and subleading large-$R$ jets in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  We see that the background distribution
  falls off rather steeply than the signal at large $p_T$
  for the two Higgs candidates.
  %
  On the other hand, the cut in $p_T$ cannot be too strong to avoid
  a substantial degradation of signal selection efficiency,
  specially taking into account the subleading large-$R$ jet.
  %
  This comparison justifies the cut of $p_T \ge 200$ GeV
  for the large-$R$ jets that we impose in the boosted category.

@@

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet must be relatively hard,
in particular they should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet, one needs that
the cut in its  $p_T$ should not bee too large, to keep most of the signal events.
%
So as in the case of the $p_T$ cut in the large-$R$ jets, also in this case there
is a compromise between suppressing backgrounds but keeping a large fraction of
signal events.

@@

Turning to the resolved category, an important aspect to account
in the corresponding selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is a consequence of the fact
that in general the boost from the Higgs decays is moderate.
%
These considerations lead to the requirement that the $p_T$ cut
for the small-$R$ jets cannot be too strong.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions are peaked at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility to being able to trigger on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges of
exploiting the resolved category in this final state.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category: as well know, the production
is mostly central (more from signal since backgrounds are dominated by
QCD $t$-channel exchange) and thus the standard rapidity cuts
are relatively harmless.

@@

One of the selection cuts with highest discrimination power is the requirement
that the invariant mass of the Higgs candidate (di)jets must lie within a window
around the nominal Higgs value.
%
In Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
cut Eq.~(\ref{higgsmasswindow})
  is applied, for the resolved and boosted categories.
%
While the signal distribution is of course peaked at the
nominal Higgs mass, the background distributions
show no particular
structure.
%
As discussed in Sect.~\ref{sec:detectormodeling}, the
width of the Higgs mass peak arises both from QCD effects
like ISR and out-of-cone radiation, as well
as from the smearing applied to final state particles.
%
While the Higgs mass window requirements
in the cut-based analysis are relatively loose, the
information of the different shape of the $m_{h}$
distribution will still be fully exploited by the MVA.
%

@@

Another important kinematical distribution of this process is the invariant mass
of the di-Higgs system.
%
This is a direct measure of the boost of the system,
which in  BSM scenarios can be substantially
enhanced, for example due to
of certain $d=6$ EFT operators~\cite{Azatov:2015oxa}.
%
One important advantage of the $b\bar{b}b\bar{b}$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching
ratios,
such as $2b2\gamma$ or $2b2\tau$.
%
With this motivation, we show in
Fig.~\ref{fig:mhh} the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved and the boosted categories.

@@

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as compared
to the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this with the MVA.
%
For the boosted category the overall trend of the $m_{hh}$ distribution
is different because of the jet selection cuts, with the
distribution now peaking at higher values of the invariant mass.
%
In this case signal and background distributions
look reasonably similar.
%
Note that at parton-level the $m_{hh}$ distribution
for signal events has a kinematical
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is modified due to both
the parton shower and to detector resolution effects.
%

Finally, in Fig.~\ref{fig:pthh} we show the transverse momentum of
the di-Higgs system, $p_T^{hh}$,
for the resolved and boosted categories.
%
Again we see that the background has a steeper $p_T^{hh}$ distribution
that the signal, in both categories, thus this variable
should provide additional discrimination power, and therefore
we will use it as another of the MVA inputs.
%
Note that in our LO simulation the $p_T^{hh}$ distribution is generated
purely
by the parton shower - a more refined treatment would require
either the matching with higher-multiplicity
matrix elements~\cite{Maierhofer:2013sha} or
the full NLO calculation~\cite{Frederix:2014hta},
%
Nevertheless, as we will show,
the MVA shows only limited sensitivity to this variable, so its
modelling appears not to be crucial in our case.

@@

\subsection{The impact of pile-up}
\label{sec:pileup}

The results shown so far did not accounted for the
effects of PU.
%
To study the impact of PU in our analysis,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios for the impact
of PU in our analysis,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts, since it
is beyond the scope of this paper to fully optimize
PU subtraction, we will adopt $\la n_{\rm PU}\ra=80$ as our baseline.

%
In order to subtract the PU, a number of techniques
have been developed
recently~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.\footnote{
These techniques have also important applications in the subtraction
of the UE/MPI contamination for jet reconstruction
in heavy ion collisions~\cite{Cacciari:2010te}.
}
%
In this work, PU  will be subtracted by means
of the {\tt SoftKiller} (SK)
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet}.
%
This recently developed method has been shown to improve~\cite{Cacciari:2014gra} on the
performance of the commonly used area-based subtraction~\cite{Cacciari:2009dp}.
%

The idea underlying {\tt SoftKiller} is based on eliminating particles
below a given cut-off in their transverse momentum, $p_T^{\rm (cut)}$, whose
value is dynamically determined in a way that makes the event-wide
transverse-momentum flow density $\rho$ vanish.
%
This $p_T$ flow density is defined as
\be
\rho\equiv{\rm median}_i \Bigg\{ \frac{p_{Ti}}{A_i}\Bigg\} \, ,
\ee
where the median is computed over all the patches $i$ with area
$A_i$ and transverse momentum $p_{Ti}$ in which the $\lp \eta,\phi\rp$ plane
is partitioned.
%
From its definition in terms of the median,
we observe that the value of $p_T^{(\rm cut)}$
will be dynamically raised until half of the patches have $\rho=0$.
%
The size and number of these patches is a free parameter of the algorithm -
here we will use square patches with length $a=0.4$.
%
We restrict ourselves to the central rapidity region,
$|\eta| \le 2.5$, for the estimation of the
$p_T$ flow density $\rho$.
%
In this analysis, {\tt SoftKiller} method is applied
to particles at the end of the parton shower, before
jet clustering.

To validate PU subtraction,
we now compare different kinematical distributions
in the case without PU and in the case
with PU subtracted with {\tt SoftKiller}.
%
We will find that
{\tt SoftKiller} exhibits a good
performance for the PU subtraction in the present case.
%
We  also obtain that the boosted category is less affected
by PU than the resolved category, as expected from the higher
$p_T$ threshold in the corresponding selection cuts.

First of all, we compare
in Fig.~\ref{fig:PUvalidation} the
invariant mass distributions of Higgs candidates for signal
events in the resolved category.
%
We show three curves: without PU,
with PU $\la n_{\rm PU}\ra=80$ but without any subtraction, and the
same but now with the SK subtraction.
%
As expected, if PU is not subtracted there is a large shift in the Higgs mass
peak, by about 40 GeV.
%
Once SK subtraction is performed, we recover a distribution much closer
to the original ones, with only a small shift of $\simeq 5$ GeV in the mass
distribution.
%
The performance of SK is even better in the boosted category,
since the effects of PU are mitigated for high $p_T$ jets.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.68\textwidth]{plots/m_htot_res_signal_PUnoSK.pdf}
    \caption{\small
    The invariant mass distributions of Higgs candidates in signal
events, in the resolved category, comparing the results without PU,
with PU $\la n_{\rm PU}\ra=80$ but without any subtraction, and the
corresponding results now with SK subtraction.
}
\label{fig:PUvalidation}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Next, we compare additional
 signal distributions with
and without PU, with $\la n_{\rm PU}\ra=80$ in the
former
case (and SK subtraction).
%
In Fig.~\ref{fig:m_H_PU} we show the invariant mass distribution
of the leading and subleading Higgs candidates, corresponding to both
the boosted
and resolved categories.
%
These distributions are plotted after the $b$-tagging, that is,
before they are used as input to the MVA.
%
As we can see, in the boosted category, the residual effects of PU
after the {\tt SoftKiller} subtraction are rather moderate,
with the position of the Higgs mass peaks essentially
unchanged, and only a moderate distortion of the
distribution found.
%
The effects are more important for the resolved category, where PU
shifts the Higgs peak by an amount $\Delta m_h \simeq 5$ GeV.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_res_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_res_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_bst_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_bst_comp.pdf}
  \caption{\small
    Comparison of the invariant mass distributions of the leading (left plots)
    and subleading (right plots) Higgs candidates in the resolved
    (upper plots) and boosted (lower plots) categories,
    both without PU and with
    PU, $\la n_{PU}\ra=80$, subtracted with {\tt SoftKiller}.
}
\label{fig:m_H_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Next we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in Fig.~\ref{fig:mHH_PU}, both for the boosted and
for the resolved categories.
%
In the case of the $p_T$ distribution, the differences between the selection
criteria for the resolved
and boosted categories is reflected in the rightward shift of the latter.
%
The effect of PU is negligible in the boosted case, and small
in the resolved case, except for large $p_T$ values.
%
For the case of the $m_{hh}$ distribution, similar conclusions
apply.
%
These comparisons validate the PU subtraction strategy
adopted in this work.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \vspace{-1cm}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_bst_comp.pdf}
  \caption{\small
    Comparison of the transverse momentum $p_T^h$ of the leading
    Higgs candidate (upper plots) and of the invariant mass $m_{hh}$
    of the di-Higgs system (lower plots) in the resolved
    (left plots) and boosted (right plots) categories,
    without PU and with $\la n_{PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:mHH_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

We can also assess the impact of PU in representative
substructure variables
used as input to the MVA in the boosted category.
%
In particular we consider the subjetiness variable,
$\tau_{21}$, Eq.~(\ref{eq:tau21}), and the ratio
of energy correlation functions, $D_2^{(\beta)}$,
Eq.~(\ref{eq:d2}),
corresponding to the leading Higgs candidate.
%
This comparison is illustrated in Fig.~\ref{fig:Substructure_PU}.
%
As can be seen, these substructure variables, which
as demonstrated before carry a substantial
discrimination power, are relatively unaffected by PU.
%
Recall that the $D_2^{(\beta)}$ variable is
explicitly constructed~\cite{Larkoski:2013eya}
to be resilient
with respect to PU contamination.
%
Therefore, we do not expect a major loss of discrimination
power due to PU effects in our analysis.
%

>>>>

For the   boosted and intermediate categories,
which involve the use of large-$R$ jets,
we will use jet substructure variables~\cite{Salam:2009jx,Aad:2013gja} to
improve the significance of the discrimination between signal and background
events in the MVA.
%
In particular we will consider the following jet
substructure variables:
\begin{itemize}
\item The $k_T$-splitting scale~\cite{Butterworth:2002tt,Butterworth:2008iy}.

  This variable is defined by reclustering the constituents of a jet with the
  $k_t$ algorithm~\cite{Ellis:1993tq},
  which usually clusters last the harder constituents, and then
  taking the $k_t$ distance measure between the two subjets at the final stage of the recombination
  procedure,
  \be
  \label{eq:ktsplitting}
\sqrt{d_{12}} \equiv {\rm min}\lp p_{T,1},p_{T,2}\rp \cdot \Delta R_{12} \, .
\ee
with $p_{T,1}$ and $p_{T,2}$ the transverse momenta of the two subjets merged
in the final step of the clustering, and $\Delta R_{12}$ the corresponding
angular separation.

It is also possible to define related variables like $\sqrt{d_{23}}$
that corresponds to the
splitting scale in the second-to-last clustering, though these
are mostly relevant for three-prong decays.

\item The ratio of 2-to-1-subjettiness $\tau_{12}$~\cite{Thaler:2010tr,Thaler:2011gf}.

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm~\cite{Catani:1993hr}
  and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  with $p_{T,k}$ is the $p_T$ of the constituent particle $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work we use as input to the MVA the ratio of 2-subjettiness to 1-subjettiness
  \be
  \label{eq:tau21}
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provides good discrimination power
  between QCD jets and jets arising from the decay of
  a heavy resonance.

\item The ratios of energy correlation functions (ECFs)  $C^{(\beta)}_2$~\cite{Larkoski:2013eya} and
  $D_2^{(\beta)}$~\cite{Larkoski:2014gra}.

  The ratio of energy correlation functions $C_2^{(\beta)}$ is defined as
  \be
  \label{eq:c2}
C_2^{(\beta)} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{\lc {\rm ECF}(2,\beta)\rc ^2} \, ,
\ee
while $D_2^{(\beta)}$ is instead defined as a double ratio of ECFs, that is
\be
e_3^{(\beta)}\equiv \frac{ {\rm ECF}(3,\beta)}{\lc {\rm ECF}(1,\beta)\rc^3} \, , \quad
  e_2^{(\beta)}\equiv \frac{ {\rm ECF}(2,\beta)}{\lc {\rm ECF}(1,\beta)\rc^2} \, , \quad
  \label{eq:d2}
D_2^{(\beta)} \equiv \frac{ e_3^{(\beta)})}{\lp e_2^{(\beta)} \rp^3} \, .
\ee
The energy correlation functions ${\rm ECF}(N,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.
  %
  The free parameter $\beta$ is set to a value of $\beta=2$,
  as recommended by the authors of Refs.~\cite{Larkoski:2013eya,Larkoski:2014gra}.
\end{itemize}

@@

In this analysis we adopt
a $b$-tagging strategy along the lines
of current ATLAS performance~\cite{Aad:2013gja,Aad:2015ydr},
though differences with respect to
the corresponding CMS
settings~\cite{Khachatryan:2011wq,Chatrchyan:2012jua}
do not modify qualitatively our results.
%
For each type of jet definition used, a different
$b$-tagging strategy is adopted:

@@

Once a Higgs boson candidate has been identified,
its invariant mass is required to lie within a fixed window
around the nominal Higgs boson mass of $m_h= 125$
GeV.
%
In particular we require the condition
\be
\label{higgsmasswindow}
|m_{h,j} - 125| < 40~{\rm GeV} \, ,\, j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of each of the two reconstructed  Higgs candidates.
%
This cut is substantially looser as the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}: the motivation
for this choice is that the MVA will take care of the optimisation of this,
as well as of other cuts that are not explicitly performed.
%
Only events where the two Higgs candidates satisfy
Eq.~(\ref{higgsmasswindow}) are classified as signal events.
%

@@

 This has been achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with higher significance turns out to be
the boosted one,
followed by the intermediate and the resolved ones, the latter two
with similar significance.
%
Therefore, for the final analysis, we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is assigned to
this category, else we check if it suits the intermediate
requirements, and failing also this, the event is attempted to
be classified in the
resolved category.
%
The resulting exclusive event samples are then separately processed
through the MVA, allowing a consistent combination
of the significance of the three event categories.

\subsection{Motivation for basic kinematic cuts}

We now motivate the values of
kinematic cuts applied to the different categories, 
comparing representative kinematic distributions between
signal and background events.
%
First to all, we present results without PU, and in the next
subsection we discuss the impact of PU
on the description of the various kinematic
distributions.
%
In the following, all
distributions will be normalized to their total integral.

%
To begin with,
in Fig.~\ref{fig:cutplots1} we show
the $p_T$ distributions
of the
  leading and subleading large-$R$ jets in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  We see that the background distribution
  falls off rather steeply than the signal at large $p_T$
  for the two Higgs candidates.
  %
  On the other hand, the cut in $p_T$ cannot be too strong to avoid
  a substantial degradation of signal selection efficiency,
  specially taking into account the subleading large-$R$ jet.
  %
  This comparison justifies the cut of $p_T \ge 200$ GeV
  for the large-$R$ jets that we impose in the boosted category.

@@

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet must be relatively hard,
in particular they should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet,
this cut in its  $p_T$
cannot be too harsh, to maintain a high signal selection
efficiency.
%
Therefore,
as for the previous distribution, also here 
a compromise between suppressing backgrounds but keeping a large fraction of
signal events is crucial.

@@

Turning to the resolved category, an important aspect to account
in the corresponding selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is a consequence of the fact
that in general the boost from the Higgs decays is moderate.
%
These considerations lead to the requirement that the $p_T$ cut
for the small-$R$ jets cannot be too strong.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions are peaked at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility to being able to trigger on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges of
exploiting the resolved category in this final state,
and hence the requirement that $p_T \ge 40$ GeV for
the small-$R$ jets.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category: as well know, the production
is mostly central (more in the case
of signal events, since backgrounds are dominated by
QCD $t$-channel exchange) and thus the standard rapidity cuts
are relatively harmless.

@@

One of the selection cuts with highest discrimination power is the requirement
that the invariant mass of the Higgs candidate (di)jets must lie within a window
around the nominal Higgs value, Eq.~(\ref{higgsmasswindow}).
%
In Fig.~\ref{fig:mHHinv} we show the invariant mass
of the leading reconstructed Higgs candidates, before the Higgs mass window
cut 
  is applied, for the resolved and boosted categories.
%
While the signal distribution is of course peaked at the
nominal Higgs mass, the background distributions
show no particular
structure.
%
The
width of the Higgs mass peak is driven both from QCD effects,
such as initial-state radiation (ISR)
and out-of-cone radiation, as well
as from the four-momenta smearing applied to final state particles.
%

@@

Another important kinematic distribution of this process is the invariant mass
of the di-Higgs system.
%
This is a direct measure of the boost of the system,
which in  BSM scenarios can be substantially
enhanced, for instance due to
of specific $d=6$ EFT operators~\cite{Azatov:2015oxa}.
%
One important advantage of the $b\bar{b}b\bar{b}$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching
ratios,
such as $2b2\gamma$ or $2b2\tau$.
%
In Fig.~\ref{fig:mhh} we show the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved and the boosted categories.

@@

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as compared
to the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this using the MVA.
%
For the boosted category the overall trend of the $m_{hh}$ distribution
is different because of the jet selection cuts, with the
distribution now peaking at higher values of the invariant mass.
%
In this case, signal and background distributions
look reasonably similar.
%
Note that at parton-level the $m_{hh}$ distribution
for signal events has a kinematic
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is smeared due
to 
the parton shower and to detector resolution effects.
%

In Fig.~\ref{fig:pthh} we show the transverse momentum of
the di-Higgs system, $p_T^{hh}$,
for the resolved and boosted categories.
%
Again we see that the background has a steeper $p_T^{hh}$ distribution
that the signal, in both categories, thus this variable
should provide additional discrimination power, and therefore
this distribution will be another of the inputs for the MVA.
%
In our LO simulation the $p_T^{hh}$ distribution is generated
by the parton shower, and and improved theoretical
description would require
either the matching with higher-multiplicity
matrix elements~\cite{Maierhofer:2013sha} or
the full NLO calculation~\cite{Frederix:2014hta},
%
Nevertheless, as we will show,
the MVA shows only limited sensitivity to this variable, so its
modelling is not crucial in this specific application.

@@

\subsection{Impact of pile-up}
\label{sec:pileup}

Now we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in a high pile-up (PU) environment.
%
To study the impact of PU,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
PU subtraction,

%
In order to subtract PU in hadronic collisions, a number of techniques
are available~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.\footnote{
These techniques have also important applications in the subtraction
of the UE/MPI contamination for jet reconstruction
in heavy ion collisions~\cite{Cacciari:2010te}.
}
%
In this work, PU  will be subtracted by means
of the {\tt SoftKiller} (SK)
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet},
whose performance has been shown to
improve the commonly used area-based subtraction~\cite{Cacciari:2009dp}.
%
The idea underlying {\tt SoftKiller} is based on eliminating particles
below a given cut-off in their transverse momentum, $p_T^{\rm (cut)}$, whose
value is dynamically determined in a way that makes the event-wide
transverse-momentum flow density $\rho$ vanish.
%
This $p_T$ flow density is defined as
\be
\rho\equiv{\rm median}_i \Bigg\{ \frac{p_{Ti}}{A_i}\Bigg\} \, ,
\ee
where the median is computed over all the patches $i$ with area
$A_i$ and transverse momentum $p_{Ti}$ in which the $\lp \eta,\phi\rp$ plane
is partitioned.
%
From its definition in terms of the median,
it follows that the value of $p_T^{(\rm cut)}$
will be dynamically raised until half of the patches have
$p_{Ti}=0$.
%
The size and number of these patches is a free parameter of the algorithm -
here we will use square patches with length $a=0.4$.
%
We restrict ourselves to the central rapidity region,
$|\eta| \le 2.5$, for the estimation of the
$p_T$ flow density $\rho$.
%
We apply {\tt SoftKiller}
to particles at the end of the parton shower, before
jet clustering.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.63\textwidth]{plots/m_htot_res_signal_PUnoSK.pdf}
    \caption{\small
    The invariant mass distributions of Higgs candidates in signal
events, in the resolved category, comparing the results without PU,
with PU $\la n_{\rm PU}\ra=80$ but without any subtraction, and the
corresponding results now with SK subtraction.
}
\label{fig:PUvalidation}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

To validate PU subtraction,
we now compare representative kinematic distributions
in the case without PU and in the case
with PU subtracted with {\tt SoftKiller}.
%
We compare first
in Fig.~\ref{fig:PUvalidation} the
invariant mass distributions of Higgs candidates for signal
events in the resolved category.
%
We show three curves: without PU,
with PU $\la n_{\rm PU}\ra=80$ but without any subtraction, and the
same but now with the SK subtraction.
%
If PU is not subtracted there is a large shift in the Higgs mass
peak, by about 40 GeV.
%
Once SK subtraction is performed, we recover a distribution much closer
to the original ones, with only a small shift of $\simeq 5$ GeV
and a broadening of the mass
distribution.
%

Next, we compare additional
 distributions with
and without PU, with $\la n_{\rm PU}\ra=80$ in the
former
case (and SK subtraction).
%
In Fig.~\ref{fig:m_H_PU} we show the invariant mass distribution
of the leading and subleading Higgs candidates,
this time corresponding
to the boosted category.
%
These distributions are plotted after the $b$-tagging, that is,
before they are used as input to the MVA.
%
As we can see, the residual effects of PU
after the {\tt SoftKiller} subtraction are smaller
in the boosted selection as compared to the resolved case,
with the position of the Higgs mass peaks essentially
unchanged, and only a moderate broadening of the
mass distribution.
%

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_bst_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_bst_comp.pdf}
  \caption{\small
    Comparison of the invariant mass distributions of the leading (left plot)
    and subleading (right plot) Higgs candidates
    in the boosted category,
    both without PU and with
    PU, $\la n_{PU}\ra=80$, subtracted with {\tt SoftKiller}.
}
\label{fig:m_H_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

In Fig.~\ref{fig:mHH_PU}
we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in both for the boosted and
for the resolved categories.
%
In the case of the $p_T^{h1}$ distribution, the differences between the selection
criteria for the resolved
and boosted categories is reflected in the rightward shift of the latter.
%
The effect of PU is small in the boosted case, and generally moderate
in the resolved case, except for large $p_T$ values.
%
Similar observation can be derived for the case of
the di-Higgs invariant mass $m_{hh}$ distribution, similar 
apply.
%
Therefore, we can consider the PU subtraction strategy
as validated for the purposes of this study, although
there could still be room for optimisation.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_bst_comp.pdf}
  \caption{\small
    Comparison of the transverse momentum $p_T^h$ of the leading
    Higgs candidate (upper plots) and of the invariant mass $m_{hh}$
    of the di-Higgs system (lower plots) in the resolved
    (left plots) and boosted (right plots) categories,
    without PU and with $\la n_{PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:mHH_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

We can also assess the impact of PU in some of
substructure variables that will be 
used as input to the MVA in the boosted category.
%
In Fig.~\ref{fig:Substructure_PU} we show the 2-to-1 subjettiness ratio
$\tau_{21}$, Eq.~(\ref{eq:tau21}), and the ratio
of energy correlation functions, $D_2^{(\beta)}$,
Eq.~(\ref{eq:d2}), in both cases
corresponding to the leading Higgs candidate.
%
Comparing the signal distributions in the cases without PU and
with PU subtracted with SK, we observe that both substructure variables
are reasonably robust in a high-PU environment.
%
The ratio of energy correlation functions $D_2^{(\beta)}$ 
was indeed constructed~\cite{Larkoski:2013eya}
to be resilient
with respect to PU contamination.
%

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/D2_h0_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/tau21_h0_bst_comp.pdf}
   \caption{\small
     Comparison of the substructure variables $D_2^{(\beta)}$ (left)
     and $\tau_{21}$ (right)
     for the leading Higgs candidate in the boosted category,
   without PU and with $\la n_{\rm PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:Substructure_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
